---
title: 'QC20202005-scovid-dataset.4'
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment = NA)

#################################################

### TODO                                      ###

#################################################

# 

#################################################

### Libraries                                 ###

#################################################


library(knitr)
library(naniar)
library(reshape2)
library(ggplot2)
library(ggthemes)
library(kableExtra)
library(magrittr)
library(mvnmle)
library(BaylorEdPsych)
library(stringr)
library(stringi)
library(visdat)

#################################################

### Functions                                 ###

#################################################

na.cutoff = 1.0

######################

### OPTIONS        ###

######################

WORKINGDIR = "H:/biostatistica/Dati/Giannini/scovid/scovid"
INFILENAME = "dset.ana.4.1300.csv"
OFILENAME = gsub(".csv", ".cleaned.csv", INFILENAME)

characterList<-NULL
dates<-NULL
factorList<-NULL
orderedList<-NULL
numericList<-NULL
integerList<-NULL
logicalList<-NULL

CHARENC         = FALSE
STRINGTRIMMING  = TRUE
BLANK2NA        = TRUE
CHAR2FACTOR     = FALSE
APPLYFACTORS    = FALSE
APPLYVARTYPE    = TRUE


######################

setVarType<-function(fileName){
  varType<-read.csv2(paste0(getwd(),"/", fileName))
  characterList<<-which(varType$type=="c")
  dates<<-which(varType$type=="d")
  factorList<<-which(varType$type=="f")
  orderedList<<-which(varType$type=="o")
  numericList<<-which(varType$type=="n")
  integerList<<-which(varType$type=="i")
  logicalList<<-which(varType$type=="l")
}

do.panel.quant<-function(d,xx,s){
  if(str_length(s>50)) { s<-substr(s, 0, 50) }
  if(check.na.limit(d[,xx], FALSE)){
    quant.var.panel<-as.matrix(cbind(min(d[,xx], na.rm = TRUE), t(quantile(d[,xx],na.rm = TRUE))[2], median(d[,xx], na.rm = TRUE), mean(d[,xx], na.rm = TRUE), sd(d[,xx], na.rm = TRUE),  t(quantile(d[,xx],na.rm = TRUE))[4], max(d[,xx],na.rm = TRUE)))
  colnames(quant.var.panel)<-c("Min", "Q2", "Median", "Mean", "SD", "Q3", "Max")
  print(kable(quant.var.panel, "html", longtable = T, booktabs = F, caption=paste0("Variable: ", s))%>%
    kable_styling(full_width = T, latex_options = c("hold_position")))
  }
}

do.plot.quant<-function(d,xx,s){
  if(check.na.limit(d[,xx], FALSE)){
    print(ggplot(d, 
        aes(x=1, y=d[,xx]))+
        geom_boxplot()+
        theme(axis.title.x=element_blank(),
              axis.title.y = element_text(size=8),
              axis.text.x=element_blank(),
              axis.ticks.x =element_blank(),
              panel.border = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_rect(fill = "grey98"),
              plot.background = element_rect(fill = "lightblue"))+
        labs(y = s))
    return(1)
  } else return(0)
}

do.panel.qual<-function(d,xx,s){
  if(str_length(s>50)) { s<-substr(s, 0, 50) }
  if(check.na.limit(d[,xx], FALSE)){
    qual.var.panel<-as.data.frame(table(d[,xx]))
    percent<-qual.var.panel[,2]/sum(qual.var.panel[,2])
    qual.var.panel<-cbind(qual.var.panel, percent)
    colnames(qual.var.panel)<-c("Levels", "Counts", "%")
    print(kable(qual.var.panel, "html", longtable = T, booktabs = T, caption=paste0("Variable: ", s))%>%
    kable_styling(full_width = T, latex_options = c("repeat_header", "hold_position"))%>%
      column_spec(1, width = "8cm"))
  }
}

do.plot.qual<-function(d,xx,s){
  if(check.na.limit(d[,xx], FALSE)){
    
    if(length(levels(as.factor(xx)))<=6){
      scaledLabels<-c(1:length(levels(as.factor(d[,xx]))))
    } else {
      scaledLabels<-""
    }
    
    print(ggplot(data=subset(d, !is.na(d[,xx])), aes(na.omit(d[,xx])))+
        geom_bar()+
        theme(axis.title.x=element_text(size=8),
              axis.title.y = element_text(size=8),
              axis.text.x=element_blank(),
              axis.ticks.x =element_blank(),
              panel.border = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_rect(fill = "grey98"),
              plot.background = element_rect(fill = "lightblue"))+
              #scale_x_discrete(labels=scaledLabels) +
              labs(y = s, x="Levels"))
    return(1)  
    } else return(0)
}

check.var.type<-function(d,x,s,p){
  r=0
  if(class(d[,x])=="integer" || class(d[,x])=="numeric") {
    if(p==FALSE) {do.panel.quant(d,x,s)}
    else {
      r=r+do.plot.quant(d,x,s)} 
  }
  else {
    if(p==FALSE) {do.panel.qual(d,x,s)}
    else {
      r=r+do.plot.qual(d,x,s)  
    }
  } 
  return(r)
}

check.na.limit<-function(x, s){
  v = round(length(which(is.na(x)))/length(x), 3)
  if(s==TRUE) print(v+"\n")
  if(v<=na.cutoff) return (TRUE)
  else return (FALSE)
}

# String encoding
encToUTF8<-function(x){
  if(class(x)=="character") {
    x<-stri_enc_toutf8(x, is_unknown_8bit = TRUE, validate = TRUE)
  }
  return(x)
}

# String trimming
trimStrings<-function(x){
  if(class(x)=="character") {
    x<-str_trim(x)
  }
  return(x)
}

# char to factor conversion
convertCharToFactor<-function(x, c){
  if(class(x)=="character") {
    x<-as.factor(x)
  }
  return(x)  
}

# blank to NA conversion
convertBlankToNA<-function(x){
  if(class(x)=="character") {
    x<-gsub("^$", NA, x)
  }
  return(x)
}

applyCharacter<-function(x, c){
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      x[,c[i]]<-as.character(x[,c[i]])  
    }  
  }
  return(x)
}

applyFactors<-function(x, c){
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      x[,c[i]]<-as.factor(x[,c[i]])  
    }  
  }
  return(x)
}

applyOrdered<-function(x, c){
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      x[,c[i]]<-as.ordered(x[,c[i]])  
    }  
  }
  return(x)
}

applyNumeric<-function(x, c){
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      x[,c[i]]<-as.numeric(x[,c[i]])  
    }  
  }
  return(x)
}

applyInteger<-function(x, c){
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      x[,c[i]]<-as.integer(x[,c[i]])  
    }  
  }
  return(x)
}

applyLogical<-function(x, c){
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      x[,c[i]]<-as.logical(x[,c[i]])  
    }  
  }
  return(x)
}

applyDate<-function(x, c, f){
  # f=format
  # f="%d/%m/%Y", f="%Y-%m-%d"
  if(missing(c)){
    return(x)
  } else {
    for(i in 1:length(c)){
      #x[,c[i]]<-as.Date(x[,c[i]], "%d/%m/%Y")
      x[,c[i]]<-as.Date(x[,c[i]], f)
    }  
  }
  return(x)
}

findLongFactor<-function(x){
  return(names(which(sapply(sapply(x[,sapply(x, is.factor)], levels), length)>30)))
}

is.Date <- function(x) inherits(x, 'Date')

findDates<-function(x){
  return(names(which(sapply(x, is.Date))))
}

sanitizeColNames<-function(d){
  colnames(d)<-gsub("_", ".", colnames(d))
  for(i in 1:length(colnames(d))){
    if(str_length(colnames(d)[i])>50){
      colnames(d)[i]<-paste0(substr(colnames(d)[i], 0, 50),"[...]")
    }
  }
  return(d)
}

```

# Data quality report
  
**Document ID:** QC20202005-scovid-dataset.4.1300.paz-QC
  
**Date:** `r format(Sys.time(), '%d/%m/%Y')`  
  
---
  
```{r}
#################################################

### Load dataset                              ###

#################################################

setwd(WORKINGDIR)

dset.import<-read.csv2(paste0(getwd(), "/", INFILENAME), stringsAsFactors = FALSE)
setVarType("vartype.dset.ana.4.1300.csv")

```
  
# Description  


**Dataset**

--- File name: `r INFILENAME`
  
--- Cases: `r dim(dset.import)[1]`
  
--- Variables: `r dim(dset.import)[2]`
  

# Results

## Quality check

```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}

# Column names sanitization
dset.import<-sanitizeColNames(dset.import)
cat("--- Column names sanitization: **Done**")
```
  
```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
if(CHARENC==TRUE){
  # String encoding
  for(i in 1:dim(dset.import)[2]){
    dset.import[,i]<-encToUTF8(dset.import[,i])
  } 
  cat("--- Charcters encoding: **Done**") 
} else {
    cat("--- Charcters encoding: **Skipped**")
}

```
  
```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
if(STRINGTRIMMING == TRUE){
  # String trimming
  for(i in 1:dim(dset.import)[2]){
    dset.import[,i]<-trimStrings(dset.import[,i])
  }  
  cat("--- String trimming: **Done**")
} else {
    cat("--- String trimming: **Skipped**")
  } 
```
  
```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
if(BLANK2NA == TRUE){
  # blank to NA conversion
  for(i in 1:dim(dset.import)[2]){
    dset.import[,i]<-convertBlankToNA(dset.import[,i])
  }    
  cat("--- Blank to NA conversion: **Done**")
} else {
    cat("--- Blank to NA conversion: **Skipped**")
  } 
```
  
```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
if(CHAR2FACTOR == TRUE){
  # char to factor conversion
  for(i in 1:dim(dset.import)[2]){
    dset.import[,i]<-convertCharToFactor(dset.import[,i])
  }
  cat("--- Character to factor conversion: **Done**")
} else {
    cat("--- Character to factor conversion: **Skipped**")
  }
```

```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
if(APPLYFACTORS == TRUE){
  if(length(factorList>0)){
    #apply factors to columns
    dset.import<-applyFactors(dset.import, factorList)  
    cat("--- Apply factors: **Done**")
    cat("\n\n")
  } else {
    cat("Factors list is empty")
    cat("--- Apply factors: **Skipped**")
    cat("\n\n")
  }
} else {
    cat("--- Apply factors: **Skipped**")
    cat("\n\n")
  }
```

```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), warning=FALSE}
if(APPLYVARTYPE == TRUE){
  
  cat("--- Apply var types to dataset..")
  cat("\n\n")
  
  # CHAR
  
  if(length(characterList>0)){
    dset.import<-applyCharacter(dset.import, characterList)  
    cat("--- Apply char types: **Done**")
    cat("\n\n")
  } else {
    cat("Character list is empty")
    cat("\n\n")
    cat("--- Apply character: **Skipped**")
    cat("\n\n")
  }
  
  # FACTOR
  
  if(length(factorList>0)){
    dset.import<-applyFactors(dset.import, factorList)  
    cat("--- Apply factor types: **Done**")
    cat("\n\n")
  } else {
    cat("Factors list is empty")
    cat("\n\n")
    cat("--- Apply factors: **Skipped**")
    cat("\n\n")
  }
  
  # ORDERED
  
  if(length(orderedList>0)){
    dset.import<-applyOrdered(dset.import, orderedList)  
    cat("--- Apply ordered types: **Done**")
    cat("\n\n")
  } else {
    cat("Ordered list is empty")
    cat("\n\n")
    cat("--- Apply ordered: **Skipped**")
    cat("\n\n")
  }
  
  # NUMERIC
  
  if(length(numericList>0)){
    dset.import<-applyNumeric(dset.import, numericList)  
    cat("--- Apply numeric types: **Done**")
    cat("\n\n")
  } else {
    cat("Numeric list is empty")
    cat("\n\n")
    cat("--- Apply numeric: **Skipped**")
    cat("\n\n")
  }
  
  # INTEGER
  
  if(length(integerList>0)){
    dset.import<-applyInteger(dset.import, integerList)  
    cat("--- Apply integer types: **Done**")
    cat("\n\n")
  } else {
    cat("Integer list is empty")
    cat("\n\n")
    cat("--- Apply integer: **Skipped**")
    cat("\n\n")
  }
  
  # LOGICAL
  
  if(length(logicalList>0)){
    dset.import<-applyLogical(dset.import, logicalList)  
    cat("--- Apply logical types: **Done**")
    cat("\n\n")
  } else {
    cat("Logical list is empty")
    cat("\n\n")
    cat("--- Apply logical: **Skipped**")
    cat("\n\n")
  }
  
  # DATE
  
  if(length(dates>0)){
    dset.import<-applyDate(dset.import, dates, "%Y-%m-%d")  
    cat("--- Apply date types: **Done**")
    cat("\n\n")
  } else {
    cat("Date list is empty")
    cat("\n\n")
    cat("--- Apply date: **Skipped**")
    cat("\n\n")
  }
  
} else {
    cat("--- Apply var types: **Skipped**")
    cat("\n\n")
  }
```  

```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), warning = FALSE}
vis_dat(dset.import, sort_type = FALSE, warn_large_data = FALSE)+
  theme(axis.title.x=element_text(size=10),
        axis.title.y = element_text(size=10),
        axis.text.x=element_blank()
  )+xlab("Variables")+
  ggtitle("Post processing var types")
```

```{r tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
# NA check
na.mat<-as.matrix(sapply(dset.import,function(x) sum(is.na(x))))
na.type<-as.matrix(sapply(dset.import,function(x) class(x)))
na.pc<-as.matrix(sapply(dset.import,function(x) round(length(which(is.na(x)))/length(x), 3)))
na.mat<-data.frame(na.mat, na.pc, na.type)
colnames(na.mat)<-c("NA Counts", "Missingness ratio", "Type")
```
  
--- NA counting: **Done**
  
--- Total missing values: `r sum(na.mat[,1])` (`r round(sum(na.mat[,1])/(dim(dset.import)[1]*dim(dset.import)[2])*100,3)`%)

--- Number of variables with more than `r na.cutoff*100`% missing values: `r length(which(na.mat[,2]>na.cutoff))`

--- MCAR test: `r if(dim(dset.import)[1]<=50) {LittleMCAR(dset.import)} else {cat("Number of variables > 50, test can not be performed.")}`

```{r fig.align="center", tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), warning=FALSE}
vis_miss(dset.import)+
  ggtitle("NA counts") +
  theme(axis.title.x=element_text(size=10),
              axis.title.y = element_text(size=10),
              axis.text.x=element_blank()
        )+xlab("Variables")
```

```{r results='asis', comment=NA, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
kable(na.mat, caption = "NA Counts", booktabs = T, longtable = T)%>%
  kable_styling(full_width = T, latex_options = c("striped", "hold_position", "repeat_header"))%>%
  column_spec(1, width = "8cm")
```

```{r results='asis', comment=NA, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
kable(na.mat[which(na.mat[,2]>na.cutoff),], "latex", caption = "Variables with missingness above cut off", booktabs = T, longtable = T)%>%
  kable_styling(full_width = T, latex_options = c("striped", "hold_position", "repeat_header"))%>%
  column_spec(1, width = "8cm")
```

```{r}
na.mat<-na.mat[which(na.mat[,2]<=na.cutoff),]
rn<-rownames(na.mat[which(na.mat[,2] <= na.cutoff),])
rn<-c(rn, rownames(na.mat[which(na.mat[,3]=="Date"),]))
dset.na.cleaned<-dset.import[, (names(dset.import) %in% rn)]
write.table(dset.na.cleaned, paste0(getwd(), "/",OFILENAME), sep=";", dec=",", row.names = FALSE, col.names = TRUE, quote = TRUE, na="NA")
```

--- Number of variables retained for downstream anlysis: `r dim(na.mat)[1]`
  
--- Numeric variables: `r length(which(na.mat[,3]=="numeric"))`  
  
--- Integer variables: `r length(which(na.mat[,3]=="integer"))` 
  
--- Factor variables: `r length(which(na.mat[,3]=="factor"))` 
  
--- Ordered variables: `r length(which(na.mat[,3]=="ordered"))`  
  
--- Character variables: `r length(which(na.mat[,3]=="character"))`
  
--- Logical variables: `r length(which(na.mat[,3]=="logical"))`
  
**Cleaned dataset:** `r OFILENAME`

```{r results='asis', comment=NA, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
kable(na.mat[which(na.mat[,2]<=na.cutoff),], "latex", caption = "Variables retained", booktabs = T, longtable = T)%>%
  kable_styling(full_width = T, latex_options = c("striped", "hold_position", "repeat_header"))%>%
  column_spec(1, width = "8cm")
```

```{r fig.align="center", fig.height=14, fig.width=8, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60), warning=FALSE}
gg_miss_var(dset.na.cleaned)+
  ggtitle("NA counts cleaned dataset") +
  theme(axis.text.y = element_text(size=8))
```

\newpage
## Descriptive statistics

```{r results='asis', comment=NA, warning = FALSE, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}

toBeRemoved<-which(colnames(dset.na.cleaned) %in% findLongFactor(dset.na.cleaned))

dates<-which(colnames(dset.na.cleaned) %in% findDates(dset.na.cleaned))

if(length(dates)>0){
  toBeRemoved<-c(toBeRemoved, dates)
}

if(length(toBeRemoved)>0){
  dset.print.tables<-dset.na.cleaned[, -toBeRemoved]  
} else {
  dset.print.tables<-dset.na.cleaned
}

cat("Removed from output:\n\n")
for(i in 1:length(toBeRemoved)){
  k=toBeRemoved[i]
  cat(paste0("-", colnames(dset.na.cleaned)[k], "\n\n"))
}
cat("\n\n")

for(i in 1:dim(dset.print.tables)[2]){
  if(dim(dset.print.tables[i])[1]!=0){
    s=colnames(dset.print.tables)[i]
    check.var.type(dset.print.tables, i, s, FALSE)  
  }
}
```
 
\newpage
## Data plots

```{r fig.height=1.7, fig.width=1.7, message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(blank=FALSE, width.cutoff=60)}

toBeRemoved<-which(colnames(dset.na.cleaned) %in% findLongFactor(dset.na.cleaned))

dates<-which(colnames(dset.na.cleaned) %in% findDates(dset.na.cleaned))

if(length(dates)>0){
  toBeRemoved<-c(toBeRemoved, dates)
}

if(length(toBeRemoved)>0){
  dset.print.plots<-dset.na.cleaned[, -toBeRemoved]
} else {
  dset.print.plots<-dset.na.cleaned
}

k=0
for(i in 1:dim(dset.print.plots)[2]){
  s=colnames(dset.print.plots)[i]
  k=k+check.var.type(dset.print.plots, i, s, TRUE)
  if(k%%4==0){cat("\n\n")}
}
```

```{r results='asis', warning = FALSE, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}

```

```{r results='asis', warning = FALSE, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}

```

```{r results='asis', warning = FALSE, tidy=TRUE, tidy.opts = list(blank = FALSE, width.cutoff = 60)}
par(mfrow=c(2,2))

```

\newpage
# Session

```{r echo=FALSE}
sessionInfo()
```

***

Dr Manfrini M., PhD  
Biostatistics  
Research and Innovation support Division  
  
Maria Cecilia Hospital  
Via Corriera, 1 - 48033 Cotignola (RA)  
Ph: +39 0545 217216



